A Neural Network Alternative to Convolutive Audio Models for Source Separation

### 2017-06-10 paper/paper.tex
I made some suggestions for the paper

\maketitle
%
\begin{abstract}
\end{abstract}
%
\begin{keywords}
Auto-encoders, source separation, deep learning.
\end{keywords}
%
\section{Introduction}
\label{sec:intro}
Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement. (Chandna monaural, Grais Single channel, Park Fully convolutional, Venkataramani end-to-end source). Currently, these networks are trained to learn discriminative audio models extensively. In other words, the spectrogram of the mixture is given as an input to the network. The goal of the network then, is to learn suitable time-frequency masks that separate the input spectrogram into the source and the interference componenents. Thus, the networks learn a basis decomposition that often works only for a specific source-interference pair, i.e., these models are not transferable. If the interfering signal changes, these networks have to be re-trained to learn suitable models to separate the new interfering signal in the mixture from the source. \\

A popular technique to learn transferable audio models for supervised source separation is the use of Nnon-negative matrix factorization~(NMF). Non-negative matrix factorization~(NMF) matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$. The notation $\mathbb{R}_{M \times N}^{\geq0}$ represents the set of matrices of non-negative elements of size $M \times N$. In this factorization, the basis matrix $\mathbf{W} \in \mathbb{R}_{M \times r}^{\geq{0}}$, the activation matrix $\mathbf{H} \in \mathbb{R}_{r \times N}^{\geq{0}}$ and $r$ represents the rank of the decomposition. In the case of audio signals, we apply such a factorization on audio spectrograms. In this setting, the columns of $\mathbf{W}$ begin to act as representative basis vectors for the source. The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time. As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network. We can interpret NMF as a non-negative auto-encoder in the following manner,
\begin{align}
    \text{$1^{\text{st}}$ layer:~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \\
    \text{$2^{\text{nd}}$ layer:~}\mathbf{X} &= g(\mathbf{W} \cdot \mathbf{H})
    \label{eq:nmfae}
\end{align}

Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R}^{\geq0}$ is an element-wise function that maps a real number to the space of positive real numbers. As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations. Although non-negativity of the models is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models. Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance. \\

>>>>

\maketitle
%
\begin{abstract}
\end{abstract}
%
\begin{keywords}
Auto-encoders, source separation, deep learning.
\end{keywords}
%
\section{Introduction}
\label{sec:intro}
Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement. (Chandna monaural, Grais Single channel, Park Fully convolutional, Venkataramani end-to-end source). Currently, these networks are trained to learn discriminative audio models extensively. In other words, the spectrogram of the mixture is given as an input to the network. The goal of the network then, is to learn suitable time-frequency masks that separate the input spectrogram into the source and the interference components. Thus, the networks learn a basis decomposition that often works only for a specific source-interference pair, i.e., these models are not transferable. If the interfering signal changes, these networks have to be re-trained to learn suitable models to separate the new interfering signal in the mixture from the source. \textbf{Cem: I am not sure if starting the paper with a generative vs discriminative (transferability) discussion is the right way. }  \\

A popular technique to learn transferable audio models for supervised source separation is the use of Non-negative matrix factorization~(NMF). Non-negative matrix factorization~(NMF) matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$. The notation $\mathbb{R}_{M \times N}^{\geq0}$ represents the set of matrices of non-negative elements of size $M \times N$. In this factorization, the basis matrix $\mathbf{W} \in \mathbb{R}_{M \times r}^{\geq{0}}$, the activation matrix $\mathbf{H} \in \mathbb{R}_{r \times N}^{\geq{0}}$ and $r$ represents the rank of the decomposition. In the case of audio signals, we apply NMF on audio spectrograms. In the audio setting, the columns of $\mathbf{W}$ begin to act as representative basis vectors for the source. The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time. \textbf{Cem: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network. We can interpret NMF as a non-negative auto-encoder in the following manner,
\begin{align}
    \text{$1^{\text{st}}$ layer:~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \\
    \text{$2^{\text{nd}}$ layer:~}\mathbf{X} &= g(\mathbf{W} \cdot \mathbf{H})
    \label{eq:nmfae}
\end{align}

Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R}^{\geq0}$ is an element-wise function that maps a real number to the space of positive real numbers. As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations. Although non-negativity of the models \textbf{Do you mean non-negativity of the parameters?} is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models. Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance. \\
